---
title: "ETC5513 Assignment 01"
author: "Chenuli Duasha Dias Muthukuda Arachchige"
format:
  html:
    theme: cerulean
  pdf: default
execute:
  echo: false
  warning: false
  message: false
---

```{r}
# Loading required libraries

library(tidyverse)
library(ggplot2)
library(knitr)
```

```{r}
# Load the assigned dataset

# load your dataset
data <- read_csv("data/32946503_Chenuli Muthukuda Arachchige.csv")
```

## Introduction 
This report will examine the relationship between marketing spend per customer and their corresponding purchase amounts in a retail context.Understanding whether higher marketing investment leads to increased purchases is important for improving budget efficiency and return on investment. By analysing customer-level campaign data, this study aims to determine whether increased marketing expenditure leads to higher purchasing behaviour and therefore improved sales outcomes. The findings will support better marketing decisions by providing insights for more efficient and data-driven strategies.

