---
title: "ETC5513 Assignment 01"
author: "Chenuli Duasha Dias Muthukuda Arachchige"
format:
  html:
    theme: cerulean
  pdf: default
execute:
  echo: false
  warning: false
  message: false
  output: false
---

```{r}
# Loading required libraries

library(tidyverse)
library(ggplot2)
library(knitr)
library(tinytex)
```

```{r}
# Load the assigned dataset

raw_campaign_data <- read_csv("data/32946503_Chenuli Muthukuda Arachchige.csv")
```

## Introduction 
This report examines the relationship between marketing expenditure per customer and their corresponding purchase amounts in a retail context. Understanding whether increased marketing investment leads to higher purchases is essential for improving budget efficiency and return on investment. By analysing customer-level campaign data, this study aims to assess the extent to which higher marketing spend influences purchase outcomes and improved sales outcomes. The findings are intended to inform more effective marketing decisions by providing insights that support efficient, data-driven strategies.


## Research Question
This report investigates the research question: "To what extent does marketing expenditure per customer influence purchase amounts?". 

The study analyses customer-level campaign data to evaluate the direction and strength of this relationship. The aim is to determine whether increased marketing investment leads to meaningful changes in customer purchasing behaviour.


## Dataset Introduction

This dataset contains customer-level campaign data, including key variables such as marketing_spend and purchase_amount, which are used to analyse the relationship between marketing investment and customer spending behaviour. Each observation represents an individual customer and captures their response to marketing efforts in terms of purchasing activity. 

Additional variables, such as visits and segment, were retained as they provide valuable context on customer engagement and enable comparisons across different customer groups to support more data-driven strategies. 

During the data cleaning process, variables such as notes and signup_date were removed as they were not directly relevant to the research question and provided limited additional insight into the relationship being analysed. Records with missing values in purchase_amount were excluded to ensure accurate analysis, as missing outcome values would reduce reliability, and any duplicate entries were removed so that each customer was represented only once in the dataset.

```{r}
# Cleaning the dataset 

# Check structure of the data
str(raw_campaign_data)

# Select only releavent variables 
processed_campaign_data <- raw_campaign_data |>
  select(customer_id,marketing_spend, purchase_amount, visits, segment)

# Remove NA values from purchases amount
processed_campaign_data <- processed_campaign_data |>
  filter(!is.na(purchase_amount))

# Remove any duplicates 
processed_campaign_data <- processed_campaign_data |>
distinct()

```



